AAAI Press Formatting Instructions \\for Authors Using \LaTeX{} --- A Guide

### 2021-09-16 publications/iaai21/dora_outlier_detection_iaai.tex

\subparagraph{Probability.}
The negative sampling algorithm is implemented by converting the unsupervised 
outlier ranking problem into a semi-supervised 
problem~\citep{sipple:neg-sampling20}. Negative (anomalous) 
examples are created by sampling from an expanded space defined by the minimum 
and maximum values of each dimension of the positive (normal) examples. The 
negative and positive examples are then used to train a random forest 
classifier. We use the posterior probabilities of the random forest classifier 
as outlier scores, which means that the observations with higher posterior 
probabilities are more likely to be outliers. 
The probabilistic autoencoder is a generative model consisting of an 
autoencoder trained to reconstruct input data which is interpreted
probabilistically after training using a normalizing flow on the autoencoder
latent space~\cite{bohm2020probabilistic}. 
Samples are scored as outliers using the log likelihood in the latent
distribution, the autoencoder reconstruction error, or a combination of
the two.

@@

\subsubsection{Targets in Mars rover images.}
% We initialized each algorithm with $D_s$ and requested a ranking of
% targets in $D_{s+}$.
\todo{Consider merging all three data set discussions since they are
in one table? with their own figures? The point here isn't really to
compare algs or data sets but instead to show that we CAN compare
them, using DORA.}
%
Table~\ref{tab:faw_results} compares the performance of each ranking 
algorithm by reporting the precision obtained when the number of
selections is the same as the total number of outliers.  We could not
employ RX in this case since the data dimensionality ($64 \times 64 =
4096$) exceeded the data set size.  We found that
the Isolation Forest achieved the highest precision (best outlier
detection) when allowed to select only \num{9} images.
Figure~\ref{fig:planetary_results} shows the complete (cumulative)
outlier detection performance for each algorithm when ranking
all \num{58} target images in $D_{1694-1703}$. \todo{Explain MDR the first
time it appears. }

@@

\begin{table}
  \caption{Precision at $N=n_{outlier}$ for three applications; the
  best result for each data set is in bold.}
  \label{tab:faw_results}
  \centering
  \begin{tabular}{l|cccc}
    \hline
    Algorithm & Astro. & Earth & Planetary & FMNIST \\
    \hline
    PCA               & & 0.41 & 0.44 \\
    DEMUD             & & 0.41 & 0.44 \\
    \hline
    RX                & & 0.43 & ---- \\
    LRX               & & ---- & 0.33 \\
    \hline
    Iso. Forest  & & 0.46 & \textbf{0.56} \\
    \hline
    PAE               & & 0.41 & 0.22 \\
    Neg. Sampling & & \textbf{0.49} & 0.33 \\
    \hline
    Random            & & 0.32 & 0.11 \\
    \hline
  \end{tabular}
\end{table}

>>>>

\subparagraph{Probability.}
The negative sampling algorithm is implemented by converting the unsupervised 
outlier ranking problem into a semi-supervised 
problem~\citep{sipple:neg-sampling20}. Negative (anomalous) 
examples are created by sampling from an expanded space defined by the minimum 
and maximum values of each dimension of the positive (normal) examples. The 
negative and positive examples are then used to train a random forest 
classifier. We use the posterior probabilities of the random forest classifier 
as outlier scores, which means that the observations with higher posterior 
probabilities are more likely to be outliers.
The probabilistic autoencoder is a generative model consisting of an
autoencoder trained to reconstruct input data which is interpreted
probabilistically after training using a normalizing flow on the autoencoder
latent space~\cite{bohm2020probabilistic}.
Samples are scored as outliers using the log likelihood in the latent
distribution, the autoencoder reconstruction error, or a combination of
both.

@@

\subsubsection{Targets in Mars rover images.}
% We initialized each algorithm with $D_s$ and requested a ranking of
% targets in $D_{s+}$.
\todo{Hannah: Consider merging all three data set discussions since they are
in one table? with their own figures? The point here isn't really to
compare algs or data sets but instead to show that we CAN compare
them, using DORA.}
%
Table~\ref{tab:faw_results} compares the performance of each ranking 
algorithm by reporting the precision obtained when the number of
selections is the same as the total number of outliers.  We could not
employ RX in this case since the data dimensionality ($64 \times 64 =
4096$) exceeded the data set size.  We found that
the Isolation Forest achieved the highest precision (best outlier
detection) when allowed to select only \num{9} images.
Figure~\ref{fig:planetary_results} shows the complete (cumulative)
outlier detection performance for each algorithm when ranking
all \num{58} target images in $D_{1694-1703}$. \todo{Hannah: Explain
MDR the first time it appears. }

@@

\begin{table}
  \caption{Precision at $N=n_{outlier}$ for three applications; the
  best result for each data set is in bold.}
  \label{tab:faw_results}
  \centering
  \begin{tabular}{l|cccc}
    \hline
    Algorithm & Astro. & Earth & Planetary & FMNIST \\
    \hline
    PCA               & & 0.41 & 0.44 \\
    DEMUD             & & 0.41 & 0.44 \\
    \hline
    RX                & & 0.43 & ---- \\
    LRX               & & ---- & 0.33 \\
    \hline
    Iso. Forest  & & 0.46 & \textbf{0.56} \\
    \hline
    PAE               & & 0.41 & 0.44 \\
    Neg. Sampling & & \textbf{0.49} & 0.33 \\
    \hline
    Random            & & 0.32 & 0.11 \\
    \hline
  \end{tabular}
\end{table}
